Dialect: A Compact Line Notation Based on SMILES

### 2022-03-18 manuscript.tex
Cites paper reporting reluctance to publish. closes #96
Cites den Joed. closes #95

Since 1988, a few sources have attempted to clarify or revise Weininger's 1988 description of SMILES. A 2003 book chapter by Weininger (referred to herein as "the Book Chapter") addressed some of the limitations of the paper. The Daylight Theory Manual ("the Manual"), a website maintained by SMILES's corporate sponsor, Daylight Chemical Information Systems, Inc. ("Daylight"), recapitulates much of the material in the Paper and the Book Chapter.\cite{daylightTheory}.

In 2007 a documentation effort that would become known as OpenSMILES began.\cite{openSMILES} OpenSMILES was conceived as "a non-proprietary specification for the SMILES language," and it addressed many of the points left open by previous SMILES documentation efforts. Noteworthy contributions included: the first formal grammar; many refinements around stereochemistry; and introduction of the idea of "standard form." Absent were detailed procedures for assigning and interpreting aromatic features, and a detailed procedure for computing implicit hydrogen count. OpenSMILES also left several points of syntactic and semantic ambiguity unaddressed.

>>>>

Since 1988, a few sources have attempted to clarify or revise Weininger's 1988 description of SMILES. A 2003 book chapter by Weininger (referred to herein as "the Book Chapter") addressed some of the limitations of the paper. The Daylight Theory Manual ("the Manual"), a website maintained by SMILES's corporate sponsor, Daylight Chemical Information Systems, Inc. ("Daylight"), recapitulates much of the material in the Paper and the Book Chapter.\cite{daylightTheory}. The lack of other primary literature sources is consistent with Weininger's reported disinterest in publication.\cite{gasteiger:2018}

In 2007 a documentation effort that would become known as OpenSMILES began.\cite{openSMILES} OpenSMILES was conceived as "a non-proprietary specification for the SMILES language," and it addressed many of the points left open by previous SMILES documentation efforts. Noteworthy contributions included: the first formal grammar, later later adapted;\cite{denjoed:2010} many refinements around stereochemistry; and introduction of the idea of "standard form." Absent were detailed procedures for assigning and interpreting aromatic features, and a detailed procedure for computing implicit hydrogen count. OpenSMILES also left several points of syntactic and semantic ambiguity unaddressed.
